---
title: "hunterdickey_A3"
author: "Hunter Dickey"
date: "10/6/2021"
output: html_document
---

In this assignment, I used the latest (2019) PUMS data to map the Bay Area's low income households with elderly residents (age >= 60). I chose this subset of the Bay Area population as it highlights two major factors that may contribute to and exacerbate extreme heat impacts: financial resource availability and age. 

Based on the insights we reviewed this past week in class (10/05 & 10/07) the elderly are a population group that is typically at higher risk to extreme heat-related illness than the younger populations. Moreover, I assume that lower-income communities may also be at a higher risk of experiencing extreme heat's adverse effects relative to their wealthier counterparts. Simply, I hold that this higher-risk assumption is because this subgroup has lower levels of household income, which may hinder its members' ability to purchase heat mitigation tools. 

The PUMS map highlights the distribution of low income households with elderly members in the Bay Area. This map could be useful for narrowing down which census blocks and communities have significantly higher concentrations of these at-risk households. According to my findings, there are 412612 households in the Bay Area that fall under the "low income with elderly members (>=60)" criteria that I set. With that figure in mind, to truly help these communities during episodes of extreme heat in the area, I think that this type of map would have to be much more precise and granular within the darker color/higher percentage blocks it features. At the moment, I believe that this map is a good starting point; but, that it lacks the specificity needed to mobilize significant action to help this particular sub population of Bay Area households. 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F)
```

```{r}
library(censusapi)
library(tidyverse)
library(tigris)
library(sf)
library(leaflet)

```

```{r}

Sys.setenv(CENSUS_KEY="d606b444d6d806f3ba91342926b0e6e9356b2501")

dec_population_2020 <-
  listCensusMetadata(
    name = "2020/dec/pl",
    type = "variables"
  )

smc_pop_2010 <-
  getCensus(
    name = "dec/pl",
    vintage = 2010,
    region = "block:*", 
    regionin = "state:06+county:081",
    vars = "P001001"
    ) %>% 
    transmute(
    block =
      paste0(state,county,tract,block),
    pop = P001001
    )

pums_2019_1yr <- getCensus(
  name = "acs/acs1/pums",
  vintage = 2019,
  region = "public use microdata area:*", 
  regionin = "state:06",
  vars = c(
    "SERIALNO",
    "SPORDER",
    "PWGTP",
    "WGTP",
    "NP",
    "HHL",
    "HINCP",
    "R60",
    "R65"
  )
)

```

Bay Area PUMAs
```{r}
ca_pumas <-
  pumas("CA", cb = T, progress_bar = F)

bay_county_names <-
  c(
    "Alameda",
    "Contra Costa",
    "Marin",
    "Napa",
    "San Francisco",
    "San Mateo",
    "Santa Clara",
    "Solano",
    "Sonoma"
  )

bay_counties <-
  counties("CA", cb = T, progress_bar = F) %>%
  filter(NAME %in% bay_county_names)

bay_pumas <-
  ca_pumas %>% 
  st_centroid() %>% 
  .[bay_counties, ] %>% 
  st_drop_geometry() %>% 
  left_join(ca_pumas %>% select(GEOID10)) %>% 
  st_as_sf()

bay_pums <-
  pums_2019_1yr %>% 
  mutate(
    PUMA = str_pad(public_use_microdata_area,5,"left","0")
  ) %>% 
  filter(PUMA %in% bay_pumas$PUMACE10)

```

```{r}
bay_pums_example <-
  bay_pums %>% 
  filter(!duplicated(SERIALNO)) %>% 
  mutate(
    WGTP = as.numeric(WGTP),
    lowinc_elderly_in_home= ifelse(
      (R60 >= 1) &
        (HINCP < 100000),
      WGTP,
      0
    )
  ) %>% 
  group_by(PUMA) %>% 
  summarize(
    perc_lowinc_elderly_in_home =
      sum(lowinc_elderly_in_home, na.rm =T)/sum(WGTP, na.rm = T)*100
  ) %>% 
  left_join(
    bay_pumas %>% 
      select(PUMACE10),
    by = c("PUMA" = "PUMACE10")
  ) %>% 
  st_as_sf()
```


```{r}
pums_pal <- colorNumeric(
  palette = "Oranges",
  domain = bay_pums_example$perc_lowinc_elderly_in_home
)

leaflet() %>%
  addTiles() %>% 
  addPolygons(
    data = bay_pums_example,
    fillColor = ~pums_pal(perc_lowinc_elderly_in_home),
    color = "white",
    opacity = 0.5,
    fillOpacity = 0.5,
    weight = 1,
    label = ~paste0(
      round(perc_lowinc_elderly_in_home), 
      "% Households with elderly residents making less than $100K"
    ),
    highlightOptions = highlightOptions(
      weight = 2,
      opacity = 1
    )
  ) %>% 
  addLegend(
    data = bay_pums_example,
    pal = pums_pal,
    values = ~perc_lowinc_elderly_in_home,
    title = "% Households<br>with elderly residents<br>making less than $100K"
  )

```


