---
title: "hunterdickey_A2"
author: "Hunter Dickey"
date: "10/2/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F)
```

```{r}
setwd("~/Documents/GitHub/hunterdickey.github.io")
```

Map of changes in population between 2010 and 2020 for a specific place in the Bay Area (e.g., a city like East Palo Alto).

I selected East Palo ALto.


```{r}
rm(list=ls())
#get census packs 

library(sf)
library(tidyverse)
library(tigris)
library(censusapi)
library(mapview)
library(leaflet)

#practice. fetch all ca counties 
ca_counties <- counties("CA", cb = T, progress_bar = F)

#coordinate reference system 
st_crs(ca_counties)

#EPSG is imortant. sf recognizes these as imputs. TIGRIS is always in 4269
#Google maps is in 4326 (effectivly the same)
#use st_transform (in website)
```

```{r}
#access census data 

Sys.setenv(CENSUS_KEY="d606b444d6d806f3ba91342926b0e6e9356b2501")

dec_population_2020 <-
  listCensusMetadata(
    name = "2020/dec/pl",
    type = "variables"
  )

smc_pop_2020 <-
  getCensus(
    name = "dec/pl",
    vintage = 2020,
    region = "block:*", 
    regionin = "state:06+county:081",
    vars = "P1_001N"
    ) %>% 
    transmute(
    block =
      paste0(state,county,tract,block),
    pop = P1_001N
    )



```


```{r}
# listCensusMetadata

dec_population_2010 <-
  listCensusMetadata(
    name = "2010/dec/pl",
    type = "variables"
  )

smc_pop_2010 <-
  getCensus(
    name = "dec/pl",
    vintage = 2010,
    region = "block:*", 
    regionin = "state:06+county:081",
    vars = "P001001"
    ) %>% 
    transmute(
    block =
      paste0(state,county,tract,block),
    pop = P001001
    )



 

```


```{r}

smc_blocks_2020 <- blocks("CA", "San Mateo", year = 2020, progress_bar = F)

epa_boundary_2020 <- places("CA", progress_bar = F) %>% 
  filter(NAME == "East Palo Alto")
epa_pop_2020 <- smc_pop_2020 %>% 
  left_join(smc_blocks_2020 %>% select(block = GEOID20)) %>% 
  st_as_sf() %>% 
  st_centroid() %>% 
  .[epa_boundary_2020, ] %>% 
  st_set_geometry(NULL) %>% 
  left_join(smc_blocks_2020 %>% select(block = GEOID20)) %>% 
  st_as_sf()


```

```{r}

smc_blocks_2010 <- blocks("CA", "San Mateo", year = 2010, progress_bar = F)

epa_boundary_2010 <- places("CA", progress_bar = F) %>% 
  filter(NAME == "East Palo Alto")
epa_pop_2010 <- smc_pop_2010 %>% 
  left_join(smc_blocks_2010 %>% select(block = GEOID10)) %>% 
  st_as_sf() %>% 
  st_centroid() %>% 
  .[epa_boundary_2010, ] %>% 
  st_set_geometry(NULL) %>% 
  left_join(smc_blocks_2010 %>% select(block = GEOID10)) %>% 
  st_as_sf()
```

```{r}
# Spacial Subsetting 
sum(epa_pop_2010$pop)
sum(epa_pop_2020$pop)


epa_boundaries <- places("CA", progress_bar = F) %>% 
  filter(NAME == "East Palo Alto")
#2020
#%>% 
  #st_transform(4269)

smc_epa_block <- smc_blocks_2020[epa_boundaries, ]

```

```{r}
leaflet() %>% 
  addProviderTiles(providers$CartoDB.Positron) %>% 
  addPolygons(
    data = epa_pop_2020,
    stroke = F,
    fillOpacity = 0.5
  ) %>% 
  addPolygons(
    data = epa_pop_2010,
    color = "red",
    weight = 0.75,
    fill = F
  )

```

```{r}
smc_blocks_2010 <- blocks("CA", "San Mateo", year = 2010, progress_bar = F)

epa_boundary_2010 <- places("CA", progress_bar = F) %>% 
  filter(NAME == "East Palo Alto")
epa_pop_difference <- smc_pop_2010 %>% 
  left_join(smc_blocks_2010 %>% select(block = GEOID10)) %>% 
  st_as_sf() %>% 
  st_centroid() %>% 
  .[epa_boundary_2010, ] %>% 
  st_set_geometry(NULL) %>% 
  left_join(smc_blocks_2010 %>% select(block = GEOID10)) %>% 
  st_as_sf()

epa_pop_2010$pop_difference_1020 <- epa_pop_2020$pop - epa_pop_2010$pop


epa_pop_difference$pop = epa_pop_2010$pop_difference_1020 

```

```{r}
# population change in people per acre from 2010-2020

epa_abs_dif_people_per_acre <- smc_pop_2010 %>% 
  left_join(smc_blocks_2010 %>% select(block = GEOID10)) %>% 
  st_as_sf() %>% 
  st_centroid() %>% 
  .[epa_boundary_2010, ] %>% 
  st_set_geometry(NULL) %>% 
  left_join(smc_blocks_2010 %>% select(block = GEOID10)) %>% 
  st_as_sf()

epa_abs_dif_people_per_acre
# 1689.6 acres in East Palo alto 

sum(epa_pop_difference$pop)

# 5078/1689.6

epa_abs_dif_people_per_acre$pop <- epa_pop_difference$pop

epa_abs_dif_people_per_acre$pop <- epa_abs_dif_people_per_acre$pop/1689.6


epa_abs_dif_people_per_acre

#increased roughly 3 people per acre on average across the block from the beginning to the end of the 2010-2020 period. 
```
For the followung 
The first interactive map below outlines the East Palo ALto 'place' boundaries and the East Palo Alto Blocks from 2010 and 2020.

```{r}
mapview(epa_boundary_2020, color = "red") + mapview(epa_pop_2020, zcol = "pop", fill = F) + mapview(epa_pop_2010, zcol = "pop", fill = F) 
```

This interactive map outlines the actual total population of East Palo Alto (in people) in 2010, 2020, and the change in population per place block in East Palo Alto from 2010-2020. These layers can be selected/applied/removed using the layer toggle on the left side of the map display. 

```{r}
mapview(epa_pop_difference, zcol = "pop") + mapview(epa_pop_2020, zcol = "pop") + mapview(epa_pop_2010, zcol = "pop", color = "blue")
```

This interactive map outlines the change in population (in people per acre) within East Palo ALto from 2010-2020. 

```{r}
mapview(epa_abs_dif_people_per_acre, zcol = "pop")
```

This final map overlays the 2010-2020 East Palo Alto total population difference, and the absolute population difference in people per acre from 2010-2020. 
```{r}
mapview(epa_pop_difference, zcol = "pop") + mapview(epa_abs_dif_people_per_acre, zcol = "pop")

```


Reflection on Assignment 2. 