---
title: "hunterdickey_A4"
author: "Hunter Dickey"
date: "10/16/2021"
output: html_document
---

---
title: "hunterdickey_A4"
author: "Hunter Dickey"
date: "10/14/2021"
output: html_document
---

Minimum Result Analysis: Race
“Is there disproportionate education attainment by race in San Mateo County?”

For the population 25 years and older in San Mateo county, there is a noticeable disproportionality to educational attainment. When compared to the share of total race groups by populations, the white racial group is underrepresented (at 38%) by 14% at the "less-than diploma level" when when juxtaposed with their 52% share of the population. Across the other education levels of 'high school graduate', 'some college', and 'bachelor's degree or higher', the white proportions are 50%, 54%, and 55% respectively, which is far closer to the 52% of their total population proportion. 

Similarly to the white population, Asian Americans are underrepresented in the 'no diploma', and 'diploma only' categories, while being much closer to their true population proportion, and slightly over represented in the 'some college' and 'bachelor's or higher' catagories respectively. 

On the flip side of this, American Indian and Alaska Native populations are noticeably over represented in the less than high school education level. American Indian and Alaska Natives make up 1.03% of this catagory's population while their true percentage is .39% of the population at large. 

Interestingly, the Black population proportion in San Mateo is underrepresented in the 'less than high school' category and  over represented in the 'high school graduate' and 'some college or associate's degree' levels. This is a positive trend I would infer, given America's history with African Americans; however, Alaska Natives/American Indians and the 'other' race category are still underrepresented at large in the 'college/higher education' levels and Black Americans are still underrepresented at the 'bachelor's degree and beyond' level. 

Secondary Result Analysis: Ethnicity 

The total San Mateo population proportion estimates are 13% Hispanic, and 26% White, and 61% other ethnicities. At the no diploma level, I found that the Hispanic population was severely over represented at 35%, the other ethnicities population was underrepresented by 4% and the white population was underrepresented at 7% of the population. At the diploma level, the Hispanic population was over represented by 7%, the white population was underrepresented by 5% and the other ethnicities category was underrepresented by 3%. Like in the race analysis the 'college/associates degree' level largely paralleled the general population smoothly while the bachelor level severely underrepresented the Hispanic community by 7%, while also over representing the white and other ethnicities populations by 6% and 2% respectively. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F)
```

```{r}
setwd("~/Documents/GitHub/hunterdickey.github.io")
```

```{r}
library(tidyverse)
library(censusapi)
Sys.setenv(CENSUS_KEY="d606b444d6d806f3ba91342926b0e6e9356b2501")

acs_vars_2019_5yr <-
  listCensusMetadata(
    name = "2019/acs/acs5",
    type = "variables"
  )
```

```{r}
race_categories <- c(
  "White alone",
  "Black or African American alone",
  "American Indian and Alaska Native alone",
  "Asian alone",
  "Native Hawaiian and Other Pacific Islander alone",
  "Some Other Race alone",
  "Two or more Races"
)



bay_edu_race_2019 <- 
  1:7 %>%
  map_dfr(function(x) {
    
    temp <- getCensus(
 name = "acs/acs5",
 vintage = "2019",
 region = "county:081",
 regionin = "state:06",
 vars = paste0("group(C15002",LETTERS[x],")")
) %>%
  select(ends_with("E") & -c(state,NAME)) %>%
  summarize_all(sum) %>%
  pivot_longer(
    everything(),
    names_to = "name",
    values_to = "estimate"
  ) %>%
  left_join(
    acs_vars_2019_5yr %>%
      select(name, label)
    #by = "name" 
    #or by = c("variable", "name")
  ) %>%
  separate(
    label,
    into = c(NA, NA, NA, "education"),
    sep = "!!"
  ) %>%
  select(-name) %>%
  filter(!is.na(education)) %>%
  mutate(
    race = race_categories[x]
  )
  })
```

```{r}

bay_race_total <-
  bay_edu_race_2019 %>% 
  group_by(race) %>% 
  summarize(estimate = sum(estimate)) %>% 
  mutate(education = "Total")

bay_edu_race_2019 %>% 
  group_by(education, race) %>% 
  summarize(estimate = sum(estimate)) %>% 
  rbind(bay_race_total) %>% 
  ggplot() +
  geom_bar(
    aes(
      x = education %>% factor(levels = rev(c("Total",unique(bay_edu_race_2019$education)))),
      y = estimate,
      fill = race %>% factor(levels = rev(unique(bay_edu_race_2019$race)))
    ),
    stat = "identity",
    position = "fill"
  ) +
  labs(
    x = "Educational Attainment",
    y = "Proportion of households",
    title = "Bay Area household Educational Attainment by race",
    fill = "Race of householder"
  ) +
  coord_flip() +
  theme(
    legend.position = "bottom",
    legend.direction = "vertical"
  )  +
  guides(
    fill = guide_legend(
      reverse = T
    )
  )

```

```{r}
#analysis
no_dip <- sum(bay_edu_race_2019$estimate[bay_edu_race_2019$education == "Less than high school diploma"])

#Total non-white
((sum(bay_race_total$estimate[1:6])/sum(bay_race_total$estimate))*100) %>% round()

#Total white
(sum(bay_edu_race_2019$estimate[bay_edu_race_2019$race== "White alone"])/sum(bay_edu_race_2019$estimate)*100) %>% round()

#Total alaska
(sum(bay_edu_race_2019$estimate[bay_edu_race_2019$race== "American Indian and Alaska Native alone"])/sum(bay_edu_race_2019$estimate)*100)

alas_no_dip <- sum(bay_edu_race_2019$estimate[bay_edu_race_2019$race== "American Indian and Alaska Native alone" & bay_edu_race_2019$education == "Less than high school diploma"])

uneducated_alas <- ((alas_no_dip/no_dip)*100)

# White categories

white_no_dip <- sum(bay_edu_race_2019$estimate[bay_edu_race_2019$race== "White alone" & bay_edu_race_2019$education == "Less than high school diploma"])

no_dip <- sum(bay_edu_race_2019$estimate[bay_edu_race_2019$education == "Less than high school diploma"])

uneducated_white <- ((white_no_dip/no_dip)*100) %>% round()

white_dip <- sum(bay_edu_race_2019$estimate[bay_edu_race_2019$race== "White alone" & bay_edu_race_2019$education == "High school graduate (includes equivalency)"])

dip <- sum(bay_edu_race_2019$estimate[bay_edu_race_2019$education == "High school graduate (includes equivalency)"])

hs_white <- ((white_dip/dip)*100) %>% round()

white_some_college <- sum(bay_edu_race_2019$estimate[bay_edu_race_2019$race== "White alone" & bay_edu_race_2019$education == "Some college or associate's degree"])

college<- sum(bay_edu_race_2019$estimate[bay_edu_race_2019$education == "Some college or associate's degree"])

some_college_white <- ((white_some_college/college)*100) %>% round()

white_bach_plus <- sum(bay_edu_race_2019$estimate[bay_edu_race_2019$race== "White alone" & bay_edu_race_2019$education == "Bachelor's degree or higher"])

bachelor<- sum(bay_edu_race_2019$estimate[bay_edu_race_2019$education == "Bachelor's degree or higher"])

higher_edu_white <- ((white_bach_plus/bachelor)*100) %>% round()


```


```{r}
ethn_categories <- c(
  "White alone, not hispanic or latino",
  "Hispanic or Latino",
  "Other, not White or latino"
)


bay_edu_ethn_white <- 
  8 %>%
  map_dfr(function(x) {
    
    temp <- getCensus(
 name = "acs/acs5",
 vintage = "2019",
 region = "county:081",
 regionin = "state:06",
 vars = paste0("group(C15002",LETTERS[x],")")
) %>%
  select(ends_with("E") & -c(state,NAME)) %>%
  summarize_all(sum) %>%
  pivot_longer(
    everything(),
    names_to = "name",
    values_to = "estimate"
  ) %>%
  left_join(
    acs_vars_2019_5yr %>%
      select(name, label)
    #by = "name" 
    #or by = c("variable", "name")
  ) %>%
  separate(
    label,
    into = c(NA, NA, NA, "education"),
    sep = "!!"
  ) %>%
  select(-name) %>%
  filter(!is.na(education)) %>%
  mutate(
    race = ethn_categories[1]
  )
  })

#2
bay_edu_ethn_hisp <- 
  9 %>%
  map_dfr(function(x) {
    
    temp <- getCensus(
 name = "acs/acs5",
 vintage = "2019",
 region = "county:081",
 regionin = "state:06",
 vars = paste0("group(C15002",LETTERS[x],")")
) %>%
  select(ends_with("E") & -c(state,NAME)) %>%
  summarize_all(sum) %>%
  pivot_longer(
    everything(),
    names_to = "name",
    values_to = "estimate"
  ) %>%
  left_join(
    acs_vars_2019_5yr %>%
      select(name, label)
    #by = "name" 
    #or by = c("variable", "name")
  ) %>%
  separate(
    label,
    into = c(NA, NA, NA, "education"),
    sep = "!!"
  ) %>%
  select(-name) %>%
  filter(!is.na(education)) %>%
  mutate(
    race = ethn_categories[2]
  )
  })

#3 
bay_edu_other_ethn_2019 <- 
  1:7 %>%
  map_dfr(function(x) {
    
    temp <- getCensus(
 name = "acs/acs5",
 vintage = "2019",
 region = "county:081",
 regionin = "state:06",
 vars = paste0("group(C15002",LETTERS[x],")")
) %>%
  select(ends_with("E") & -c(state,NAME)) %>%
  summarize_all(sum) %>%
  pivot_longer(
    everything(),
    names_to = "name",
    values_to = "estimate"
  ) %>%
  left_join(
    acs_vars_2019_5yr %>%
      select(name, label)
    #by = "name" 
    #or by = c("variable", "name")
  ) %>%
  separate(
    label,
    into = c(NA, NA, NA, "education"),
    sep = "!!"
  ) %>%
  select(-name) %>%
  filter(!is.na(education)) %>%
  mutate(
    race = ethn_categories[3]
  )
  })


bay_edu_ethn_2019 <- rbind(bay_edu_ethn_white, bay_edu_ethn_hisp, bay_edu_other_ethn_2019)

```

```{r}
bay_ethn_total <-
  bay_edu_ethn_2019 %>% 
  group_by(race) %>% 
  summarize(estimate = sum(estimate)) %>% 
  mutate(education = "Total")

bay_edu_ethn_2019 %>% 
  group_by(education, race) %>% 
  summarize(estimate = sum(estimate)) %>% 
  rbind(bay_ethn_total) %>% 
  ggplot() +
  geom_bar(
    aes(
      x = education %>% factor(levels = rev(c("Total",unique(bay_edu_ethn_2019$education)))),
      y = estimate,
      fill = race %>% factor(levels = rev(unique(bay_edu_ethn_2019$race)))
    ),
    stat = "identity",
    position = "fill"
  ) +
  labs(
    x = "Educational Attainment",
    y = "Proportion of households",
    title = "Bay Area household Educational Attainment by Ethnicity",
    fill = "Ethnicity of householder"
  ) +
  coord_flip() +
  theme(
    legend.position = "bottom",
    legend.direction = "vertical"
  )  +
  guides(
    fill = guide_legend(
      reverse = T
    )
  )

```

```{r}
bay_edu_ethn_2019


#Total white
total_white_e <- (sum(bay_edu_ethn_2019$estimate[bay_edu_ethn_2019$race== "White alone, not hispanic or latino"])/sum(bay_edu_ethn_2019$estimate)*100) %>% round()

# Total Hispanic
total_hisp_e <-(sum(bay_edu_ethn_2019$estimate[bay_edu_ethn_2019$race== "Hispanic or Latino"])/sum(bay_edu_ethn_2019$estimate)*100) %>% round()

#Neither
total_other_e <- (sum(bay_edu_ethn_2019$estimate[bay_edu_ethn_2019$race== "Other, not White or latino"])/sum(bay_edu_ethn_2019$estimate)*100) %>% round()


#No diploma 

no_dip_e <- sum(bay_edu_ethn_2019$estimate[bay_edu_ethn_2019$education == "Less than high school diploma"])

#diploma
dip_e <- sum(bay_edu_ethn_2019$estimate[bay_edu_ethn_2019$education == "High school graduate (includes equivalency)"])

#college
college_e<- sum(bay_edu_ethn_2019$estimate[bay_edu_ethn_2019$education == "Some college or associate's degree"])

#bachelor 
bachelor_e<- sum(bay_edu_ethn_2019$estimate[bay_edu_ethn_2019$education == "Bachelor's degree or higher"])



#uneducated white 
white_no_dip_e <- sum(bay_edu_ethn_2019$estimate[bay_edu_ethn_2019$race== "White alone, not hispanic or latino" & bay_edu_ethn_2019$education == "Less than high school diploma"])

#hs white
white_dip_e <- sum(bay_edu_ethn_2019$estimate[bay_edu_ethn_2019$race== "White alone, not hispanic or latino" & bay_edu_ethn_2019$education == "High school graduate (includes equivalency)"])
#college white
white_college_e <- sum(bay_edu_ethn_2019$estimate[bay_edu_ethn_2019$race== "White alone, not hispanic or latino" & bay_edu_ethn_2019$education == "Some college or associate's degree"])

#bachelor white
white_bachelor_e <- sum(bay_edu_ethn_2019$estimate[bay_edu_ethn_2019$race== "White alone, not hispanic or latino" & bay_edu_ethn_2019$education =="Bachelor's degree or higher"])

uneducated_white <- ((white_no_dip_e/no_dip_e)*100) %>% round()

hs_white <- ((white_dip_e/dip_e)*100) %>% round()

college_white <- ((white_college_e/college_e)*100) %>% round()

bachelor_white <- ((white_bachelor_e/bachelor_e)*100) %>% round()


#uneducated Hispanic
hisp_no_dip_e<- sum(bay_edu_ethn_2019$estimate[bay_edu_ethn_2019$race== "Hispanic or Latino" & bay_edu_ethn_2019$education == "Less than high school diploma"])

#hs hispanic 
hisp_dip_e <- sum(bay_edu_ethn_2019$estimate[bay_edu_ethn_2019$race== "Hispanic or Latino" & bay_edu_ethn_2019$education == "High school graduate (includes equivalency)"])

#college hispanic
hisp_college_e <- sum(bay_edu_ethn_2019$estimate[bay_edu_ethn_2019$race== "Hispanic or Latino" & bay_edu_ethn_2019$education == "Some college or associate's degree"])

#bachelor hispanic
hisp_bachelor_e <- sum(bay_edu_ethn_2019$estimate[bay_edu_ethn_2019$race== "Hispanic or Latino" & bay_edu_ethn_2019$education =="Bachelor's degree or higher"])

uneducated_hisp <- ((hisp_no_dip_e/no_dip_e)*100) %>% round()

hs_hisp <- ((hisp_dip_e/dip_e)*100) %>% round()

college_hisp <- ((hisp_college_e/college_e)*100) %>% round()

bachelor_hisp <- ((hisp_bachelor_e/bachelor_e)*100) %>% round()


#uneducated Other
other_no_dip_e <- sum(bay_edu_ethn_2019$estimate[bay_edu_ethn_2019$race== "Other, not White or latino" & bay_edu_ethn_2019$education == "Less than high school diploma"])

#hs Other
other_dip_e <- sum(bay_edu_ethn_2019$estimate[bay_edu_ethn_2019$race== "Other, not White or latino" & bay_edu_ethn_2019$education == "High school graduate (includes equivalency)"])

#college Other
other_college_e <- sum(bay_edu_ethn_2019$estimate[bay_edu_ethn_2019$race== "Other, not White or latino" & bay_edu_ethn_2019$education == "Some college or associate's degree"])

#bachelor Other
other_bachelor_e <- sum(bay_edu_ethn_2019$estimate[bay_edu_ethn_2019$race== "Other, not White or latino" & bay_edu_ethn_2019$education =="Bachelor's degree or higher"])


uneducated_other <- ((other_no_dip_e/no_dip_e)*100) %>% round()

hs_other <- ((other_dip_e/dip_e)*100) %>% round()

college_other <- ((other_college_e/college_e)*100) %>% round()

bachelor_other <- ((other_bachelor_e/bachelor_e)*100) %>% round()



```

