---
title: "Working."
author: "Hunter Dickey"
date: "12/2/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



What factors could make people vulnerable to wildfires: household income HINCP,Fire Insurance INSP, running water RWAT, SMARTPHONE, Telephone service TEL, PUMA, VEH Vehicle available, Health insurance HICOV, POPPIV +/- RAC1P



household income HINCP
Fire Insurance INSP
Health insurance HICOV
Vehicle available VEH
Poverty ratio POPPIV

K202002
K202302
K202701
B08014
B08015

B08105A (means of transport)
B17020A Poverty 
B19001A HINCP 
C27001A Health 

```{r}

Sys.setenv(CENSUS_KEY="d606b444d6d806f3ba91342926b0e6e9356b2501")
acs_vars_2019_1yr <-
  listCensusMetadata(
    name = "2019/acs/acs1",
    type = "variables"
  )

```

  
```{r}
library(censusapi)
library(tidyverse)
library(tigris)
library(sf)
library(leaflet)
library(mapview)
```

Variable by variable Race Composition 
```{r}
race_categories <- c(
  "White alone",
  "Black or African American alone",
  "American Indian and Alaska Native alone",
  "Asian alone",
  "Native Hawaiian and Other Pacific Islander alone",
  "Some Other Race alone",
  "Two or more Races"
)

race_categories <- c(
  "White alone",
  "Black or African American alone",
  "American Indian and Alaska Native alone",
  "Asian alone",
  "Native Hawaiian and Other Pacific Islander alone",
  "Some Other Race alone",
  "Two or more Races"
)

#income disparity 
bay_income_race_2019 <- 
  1:7 %>%
  map_dfr(function(x) {
    
    temp <- getCensus(
 name = "acs/acs5",
 vintage = "2019",
 region = "county:001,013,041,055,075,081,085,095,097",
 regionin = "state:06",
 vars = paste0("group(B19001",LETTERS[x],")")
) %>%
  select(ends_with("E") & -c(state,NAME)) %>%
  summarize_all(sum) %>%
  pivot_longer(
    everything(),
    names_to = "name",
    values_to = "estimate"
  ) %>%
  left_join(
    acs_vars_2019_5yr %>%
      select(name, label)
    #by = "name" 
    #or by = c("variable", "name")
  ) %>%
  separate(
    label,
    into = c(NA, NA, "income"),
    sep = "!!"
  ) %>%
  select(-name) %>%
  filter(!is.na(income)) %>%
  mutate(
    race = race_categories[x]
  )
  })


bay_race_total <-
  bay_income_race_2019 %>% 
  group_by(race) %>% 
  summarize(estimate = sum(estimate)) %>% 
  mutate(income = "Total")

bay_income_race_2019 %>% 
  group_by(income, race) %>% 
  summarize(estimate = sum(estimate)) %>% 
  rbind(bay_race_total) %>% 
  ggplot() +
  geom_bar(
    aes(
      x = income %>% factor(levels = rev(c("Total",unique(bay_income_race_2019$income)))),
      y = estimate,
      fill = race %>% factor(levels = rev(unique(bay_income_race_2019$race)))
    ),
    stat = "identity",
    position = "fill"
  ) +
  labs(
    x = "Household income",
    y = "Proportion of households",
    title = "Bay Area household income by race",
    fill = "Race of householder"
  ) +
  coord_flip() +
  theme(
    legend.position = "bottom",
    legend.direction = "vertical"
  )  +
  guides(
    fill = guide_legend(
      reverse = T
    )
  )
```

Health Insurance disparity 
```{r}

race_categories <- c(
  "White alone",
  "Black or African American alone",
  "American Indian and Alaska Native alone",
  "Asian alone",
  "Native Hawaiian and Other Pacific Islander alone",
  "Some Other Race alone",
  "Two or more Races"
)

#health disparity 
bay_healthin_race_2019 <- 
  1:7 %>%
  map_dfr(function(x) {
    
    temp <- getCensus(
 name = "acs/acs5",
 vintage = "2019",
 region = "county:001,013,041,055,075,081,085,095,097",
 regionin = "state:06",
 vars = paste0("group(C27001",LETTERS[x],")")
) %>%
  select(ends_with("E") & -c(state,NAME)) %>%
  summarize_all(sum) %>%
  pivot_longer(
    everything(),
    names_to = "name",
    values_to = "estimate"
  ) %>%
  left_join(
    acs_vars_2019_5yr %>%
      select(name, label)
    #by = "name" 
    #or by = c("variable", "name")
  ) %>%
  separate(
    label,
    into = c(NA, NA, NA, "HealthIn"),
    sep = "!!"
  ) %>%
  select(-name) %>%
  filter(!is.na(HealthIn)) %>%
  mutate(
    race = race_categories[x]
  )
  })


bay_race_total <-
  bay_healthin_race_2019 %>% 
  group_by(race) %>% 
  summarize(estimate = sum(estimate)) %>% 
  mutate(income = "Total")

bay_healthin_race_2019 %>% 
  group_by(HealthIn, race) %>% 
  summarize(estimate = sum(estimate)) %>% 
  rbind(bay_race_total) %>% 
  ggplot() +
  geom_bar(
    aes(
      x = HealthIn %>% factor(levels = rev(c("Total",unique(bay_healthin_race_2019$HealthIn)))),
      y = estimate,
      fill = race %>% factor(levels = rev(unique(bay_healthin_race_2019$race)))
    ),
    stat = "identity",
    position = "fill"
  ) +
  labs(
    x = "Health Insurance Coverage",
    y = "Proportion of households",
    title = "Bay Area household Health Coverage by race",
    fill = "Race of householder"
  ) +
  coord_flip() +
  theme(
    legend.position = "bottom",
    legend.direction = "vertical"
  )  +
  guides(
    fill = guide_legend(
      reverse = T
    )
  )
```

Transportation
```{r}
bay_transportation_race_2019 <- 
  1:7 %>%
  map_dfr(function(x) {
    
    temp <- getCensus(
 name = "acs/acs5",
 vintage = "2019",
 region = "county:001,013,041,055,075,081,085,095,097",
 regionin = "state:06",
 vars = paste0("group(B08105",LETTERS[x],")")
) %>%
  select(ends_with("E") & -c(state,NAME)) %>%
  summarize_all(sum) %>%
  pivot_longer(
    everything(),
    names_to = "name",
    values_to = "estimate"
  ) %>%
  left_join(
    acs_vars_2019_5yr %>%
      select(name, label)
    #by = "name" 
    #or by = c("variable", "name")
  ) %>%
  separate(
    label,
    into = c(NA, NA, "Transportation"),
    sep = "!!"
  ) %>%
  select(-name) %>%
  filter(!is.na(Transportation)) %>%
  mutate(
    race = race_categories[x]
  )
  })


bay_race_total <-
  bay_transportation_race_2019 %>% 
  group_by(race) %>% 
  summarize(estimate = sum(estimate)) %>% 
  mutate(Transportation = "Total")

bay_transportation_race_2019 %>% 
  group_by(Transportation, race) %>% 
  summarize(estimate = sum(estimate)) %>% 
  rbind(bay_race_total) %>% 
  ggplot() +
  geom_bar(
    aes(
      x = Transportation %>% factor(levels = rev(c("Total",unique(bay_transportation_race_2019$Transportation)))),
      y = estimate,
      fill = race %>% factor(levels = rev(unique(bay_transportation_race_2019$race)))
    ),
    stat = "identity",
    position = "fill"
  ) +
  labs(
    x = "Household Transportation Method",
    y = "Proportion of households",
    title = "Bay Area household transportation method by race",
    fill = "Race of householder"
  ) +
  coord_flip() +
  theme(
    legend.position = "bottom",
    legend.direction = "vertical"
  )  +
  guides(
    fill = guide_legend(
      reverse = T
    )
  )

```


Poverty Status Indicator 
```{r}
bay_poverty_race_2019 <- 
  1:7 %>%
  map_dfr(function(x) {
    
    temp <- getCensus(
 name = "acs/acs5",
 vintage = "2019",
 region = "county:001,013,041,055,075,081,085,095,097",
 regionin = "state:06",
 vars = paste0("group(B17020",LETTERS[x],")")
) %>%
  select(ends_with("E") & -c(state,NAME)) %>%
  summarize_all(sum) %>%
  pivot_longer(
    everything(),
    names_to = "name",
    values_to = "estimate"
  ) %>%
  left_join(
    acs_vars_2019_5yr %>%
      select(name, label)
    #by = "name" 
    #or by = c("variable", "name")
  ) %>%
  separate(
    label,
    into = c(NA, NA, "poverty"),
    sep = "!!"
  ) %>%
  select(-name) %>%
  filter(!is.na(poverty)) %>%
  mutate(
    race = race_categories[x]
  )
  })


bay_race_total <-
  bay_poverty_race_2019 %>% 
  group_by(race) %>% 
  summarize(estimate = sum(estimate)) %>% 
  mutate(poverty = "Total")

bay_poverty_race_2019 %>% 
  group_by(poverty, race) %>% 
  summarize(estimate = sum(estimate)) %>% 
  rbind(bay_race_total) %>% 
  ggplot() +
  geom_bar(
    aes(
      x = poverty %>% factor(levels = rev(c("Total",unique(bay_poverty_race_2019$poverty)))),
      y = estimate,
      fill = race %>% factor(levels = rev(unique(bay_poverty_race_2019$race)))
    ),
    stat = "identity",
    position = "fill"
  ) +
  labs(
    x = "Household income above or below the poverty line",
    y = "Proportion of households",
    title = "Bay Area household poverty proportion by race",
    fill = "Race of householder"
  ) +
  coord_flip() +
  theme(
    legend.position = "bottom",
    legend.direction = "vertical"
  )  +
  guides(
    fill = guide_legend(
      reverse = T
    )
  )
```


#########################################################

Distribution (mapping these disparities, pums)
```{r}
pums_2019_1yr <- readRDS("project_pums.rds")

ca_pumas <-
  pumas("CA", cb = T, progress_bar = F)

bay_county_names <-
  c(
    "Alameda",
    "Contra Costa",
    "Marin",
    "Napa",
    "San Francisco",
    "San Mateo",
    "Santa Clara",
    "Solano",
    "Sonoma"
  )

bay_counties <-
  counties("CA", cb = T, progress_bar = F) %>%
  filter(NAME %in% bay_county_names)

bay_pumas <-
  ca_pumas %>% 
  st_centroid() %>% 
  .[bay_counties, ] %>% 
  st_drop_geometry() %>% 
  left_join(ca_pumas %>% select(GEOID10)) %>% 
  st_as_sf()

bay_pums <-
  pums_2019_1yr %>% 
  mutate(
    PUMA = str_pad(public_use_microdata_area,5,"left","0")
  ) %>% 
  filter(PUMA %in% bay_pumas$PUMACE10)

```

regions low income poverty line, 
```{r}
bay_pums_low_income <-
  bay_pums %>% 
  filter(!duplicated(SERIALNO)) %>% 
  mutate(
    WGTP = as.numeric(WGTP),
    poverty= ifelse(
      (POVPIP <= 0),
      WGTP,
      0
    )
  ) %>% 
  group_by(PUMA) %>% 
  summarize(
    poverty =
      sum(poverty, na.rm =T)/sum(WGTP, na.rm = T)*100
  ) %>% 
  left_join(
    bay_pumas %>% 
      select(PUMACE10),
    by = c("PUMA" = "PUMACE10")
  ) %>% 
  st_as_sf()
```
Leaflet
```{r}
pums_pal <- colorNumeric(
  palette = "Purples",
  domain = bay_pums_low_income$poverty
)

leaflet() %>%
  addTiles() %>% 
  addPolygons(
    data = bay_pums_low_income,
    fillColor = ~pums_pal(poverty),
    color = "white",
    opacity = 0.5,
    fillOpacity = 0.5,
    weight = 1,
    label = ~paste0(
      round(poverty), 
      "% Households at or below the poverty line"
    ),
    highlightOptions = highlightOptions(
      weight = 2,
      opacity = 1
    )
  ) %>% 
  addLegend(
    data = bay_pums_low_income,
    pal = pums_pal,
    values = ~poverty,
    title = "% Households at or <br>below the<br>poverty line"
  )

```

health 

```{r}

as.numeric(pums_2019_1yr$HICOV)



bay_pums_lacking_health_coverage <-
  bay_pums %>% 
  filter(!duplicated(SERIALNO)) %>% 
  mutate(
    WGTP = as.numeric(WGTP),
    no_health= ifelse(
      (HICOV >=2),
      WGTP,
      0
    )
  ) %>% 
  group_by(PUMA) %>% 
  summarize(
    no_health =
      sum(no_health, na.rm =T)/sum(WGTP, na.rm = T)*100
  ) %>% 
  left_join(
    bay_pumas %>% 
      select(PUMACE10),
    by = c("PUMA" = "PUMACE10")
  ) %>% 
  st_as_sf()
```
Leaflet
```{r}
pums_pal <- colorNumeric(
  palette = "Reds",
  domain = bay_pums_lacking_health_coverage$no_health
)

leaflet() %>%
  addTiles() %>% 
  addPolygons(
    data = bay_pums_lacking_health_coverage,
    fillColor = ~pums_pal(no_health),
    color = "white",
    opacity = 0.5,
    fillOpacity = 0.5,
    weight = 1,
    label = ~paste0(
      round(no_health), 
      "% Households lacking Health Insurance"
    ),
    highlightOptions = highlightOptions(
      weight = 2,
      opacity = 1
    )
  ) %>% 
  addLegend(
    data = bay_pums_lacking_health_coverage,
    pal = pums_pal,
    values = ~no_health,
    title = "% Households lacking <br>Health<br>Insurance"
  )
```

transportation 

```{r}
bay_pums_no_vehicle <-
  bay_pums %>% 
  filter(!duplicated(SERIALNO)) %>% 
  mutate(
    WGTP = as.numeric(WGTP),
    no_car= ifelse(
      (VEH < 1),
      WGTP,
      0
    )
  ) %>% 
  group_by(PUMA) %>% 
  summarize(
    no_car =
      sum(no_car, na.rm =T)/sum(WGTP, na.rm = T)*100
  ) %>% 
  left_join(
    bay_pumas %>% 
      select(PUMACE10),
    by = c("PUMA" = "PUMACE10")
  ) %>% 
  st_as_sf()
```
Leaflet
```{r}
pums_pal <- colorNumeric(
  palette = "Oranges",
  domain = bay_pums_no_vehicle$no_car
)

leaflet() %>%
  addTiles() %>% 
  addPolygons(
    data = bay_pums_no_vehicle,
    fillColor = ~pums_pal(no_car),
    color = "white",
    opacity = 0.5,
    fillOpacity = 0.5,
    weight = 1,
    label = ~paste0(
      round(no_car), 
      "% Households Without an Accessible Vehicle"
    ),
    highlightOptions = highlightOptions(
      weight = 2,
      opacity = 1
    )
  ) %>% 
  addLegend(
    data = bay_pums_no_vehicle,
    pal = pums_pal,
    values = ~no_car,
    title = "% Households Without <br>an Accessible<br>Vehicle"
  )
```

Fire Coverage below average/sufficient coverage
```{r}
bay_pums_fire_coverage <-
  bay_pums %>% 
  filter(!duplicated(SERIALNO)) %>% 
  mutate(
    WGTP = as.numeric(WGTP),
    with_fire_coverage= ifelse(
      (INSP > 1000),
      WGTP,
      0
    )
  ) %>% 
  group_by(PUMA) %>% 
  summarize(
    with_fire_coverage =
      sum(with_fire_coverage, na.rm =T)/sum(WGTP, na.rm = T)*100
  ) %>% 
  left_join(
    bay_pumas %>% 
      select(PUMACE10),
    by = c("PUMA" = "PUMACE10")
  ) %>% 
  st_as_sf()

```
Leaflet
```{r}
pums_pal <- colorNumeric(
  palette = "Oranges",
  domain = bay_pums_fire_coverage$with_fire_coverage
)

leaflet() %>%
  addTiles() %>% 
  addPolygons(
    data = bay_pums_fire_coverage,
    fillColor = ~pums_pal(with_fire_coverage),
    color = "white",
    opacity = 0.5,
    fillOpacity = 0.5,
    weight = 1,
    label = ~paste0(
      round(with_fire_coverage), 
      "% Households insured For Fire"
    ),
    highlightOptions = highlightOptions(
      weight = 2,
      opacity = 1
    )
  ) %>% 
  addLegend(
    data = bay_pums_fire_coverage,
    pal = pums_pal,
    values = ~with_fire_coverage,
    title = "% Households<br>insured<br>for Fire"
  )
```
###
Fire High Risk vs Occurance Map
```{r}

library(tidyverse)
library(sf)
library(cowplot)
library(raster)
library(scales)

knitr::opts_knit$set(message = F, warning = F)

whp <- raster("~/Documents/GitHub/hunterdickey.github.io/fire-maps/whp2020_GeoTIF/whp2020_cnt_conus.tif")
```

```{r}
bay_county_names <-
  c(
    "Alameda",
    "Contra Costa",
    "Marin",
    "Napa",
    "San Francisco",
    "San Mateo",
    "Santa Clara",
    "Solano",
    "Sonoma"
  )
bay_area_fire_df <- 1:9 %>% map_dfr(function(x) {
  bay_area_fire_santa_clara <- st_read("~/Documents/GitHub/hunterdickey.github.io/fire-maps/cb_2015_us_county_20m", quiet = T) %>%
  filter(NAME == bay_county_names[x]) %>%
  st_transform(proj4string(whp))
})


bay_area_fire_santa_clara <- st_read("~/Documents/GitHub/hunterdickey.github.io/fire-maps/cb_2015_us_county_20m", quiet = T) %>%
  filter(NAME == "Santa Clara") %>%
  st_transform(proj4string(whp))


whp <- crop(whp, bay_area_fire_df)


whp_df <- as.data.frame(whp, xy=TRUE)

ggplot(bay_area_fire_df) + 
  geom_raster(data = whp_df, mapping = aes(x = x, y = y, fill = whp2020_cnt_conus)) + 
  geom_sf(fill = NA) + 
  scale_fill_distiller(palette = "PuBuGn", direction = 1) + 
  guides(fill = "none") + 
  theme_map()

```

Fire Perimeters
```{r}
setwd("~/Documents/GitHub/hunterdickey.github.io")
perims <- st_read("~/Documents/GitHub/hunterdickey.github.io/fire-maps/fire20_1.gdb/", layer = "firep20_1")

perims_bay <- perims %>% 
  st_cast("MULTIPOLYGON") %>% # Avoid Unknown WKB type 12 error
  st_make_valid() %>% # Fixes other possible errors
  st_transform(st_crs(bay_area_fire_df)) %>% 
  st_crop(bay_area_fire_df)
```

Credit: 
Patrick Baylis's Research on wildfire mapping
```{r}
ggplot(bay_area_fire_df) + 
  geom_raster(data = whp_df, mapping = aes(x = x, y = y, fill = whp2020_cnt_conus)) + 
  geom_sf(data = perims_bay, fill = "Orange", colour = NA, alpha = 0.25) + 
  geom_sf(fill = NA) + 
  scale_fill_distiller(palette = "PuBuGn", direction = 1) + 
  guides(fill = FALSE) + 
  theme_map()
```

Logistical Regression : risk 
```{r}
pums_2019_1yr <- readRDS("project_pums.rds")

```

```{r}
ca_pumas <-
  pumas("CA", cb = T, progress_bar = F)

bay_counties <-
  counties("CA", cb = T, progress_bar = F) %>%
  filter(NAME %in% bay_county_names)

bay_pumas <-
  ca_pumas %>% 
  st_centroid() %>% 
  .[bay_counties, ] %>% 
  st_drop_geometry() %>% 
  left_join(ca_pumas %>% select(GEOID10)) %>% 
  st_as_sf()

bay_pums <-
  pums_2019_1yr %>% 
  mutate(
    PUMA = str_pad(public_use_microdata_area,5,"left","0")
  ) %>% 
  filter(PUMA %in% bay_pumas$PUMACE10)


```

```{r}
mapview(bay_pumas)
```

```{r}
bay_pums_clean <- bay_pums %>% 
  mutate(
    HINCP = as.numeric(HINCP),
    HICOV = as.numeric(HICOV),
    POVPIP = as.numeric(POVPIP),
    VEH = as.numeric(VEH),
    firerisk = ifelse(
      as.numeric(POVPIP) <= 0 &
        as.numeric(HINCP) < 90000 &
        as.numeric(HICOV)>=2, 
      1, 
      0
      ),
  ) %>%
  arrange(RAC1P) %>%
  group_by(SERIALNO) %>%
  summarize_all(first)



# B08105A (means of transport)
# B17020A Poverty 
# B19001A HINCP 
# C27001A Health 
```


LOGIT MODEL

```{r}
bay_pums_clean <- bay_pums_clean %>% 
  mutate(
    race = RAC1P %>% 
      factor(
        levels = bay_pums_clean$RAC1P %>% 
          unique() %>%
          as.numeric() %>% 
          sort()
        ),
    health_coverage = HICOV %>% 
      factor(
        levels = bay_pums_clean$HICOV %>% 
          unique() %>%
          as.numeric() %>% 
          sort()
        ),
    income = HINCP %>% 
      factor(
        levels = bay_pums_clean$HINCP %>% 
          unique() %>%
          as.numeric() %>% 
          sort()
        ),
    vehicle = VEH %>% 
      factor(
        levels = bay_pums_clean$VEH %>% 
          unique() %>%
          as.numeric() %>% 
          sort()
      ),
     PUMA = as.factor(PUMA)
  )

```

```{r}
logit_model <- glm(
  firerisk ~ race + health_coverage + vehicle + PUMA,
  family = quasibinomial(),
  data = bay_pums_clean
)

print(summary(logit_model))

```










Logistical Regression : risk 
```{r}
pums_2019_1yr <- readRDS("project_pums.rds")

```

```{r}
library(tidyverse)
library(sf)
library(cowplot)
library(raster)
library(scales)
library(censusapi)
library(tigris)
library(leaflet)
library(mapview)
ca_pumas <-
  pumas("CA", cb = T, progress_bar = F)

 
sf_boundary <-
  counties("CA", cb = T, progress_bar = F) %>%
  filter(NAME == "San Francisco")

sf_pumas <-
  ca_pumas %>% 
  st_centroid() %>% 
  .[sf_boundary, ] %>% 
  st_drop_geometry() %>% 
  left_join(ca_pumas %>% select(GEOID10)) %>% 
  st_as_sf()

sf_pums <- pums_2019_1yr %>% 
  mutate(
    PUMA = str_pad(public_use_microdata_area,5,"left","0")
  ) %>%
  filter(PUMA %in% sf_pumas$PUMACE10)


```

```{r}
mapview(sf_pumas)
```

```{r}
sff_pums_clean <- sf_pums %>% 
  mutate(
    HINCP = as.numeric(HINCP),
    HICOV = as.numeric(HICOV),
    POVPIP = as.numeric(POVPIP),
    VEH = as.numeric(VEH),
    firerisk = ifelse(
      as.numeric(POVPIP) <= 0 &
        as.numeric(HINCP) < 90000 &
        as.numeric(HICOV)>=2, 
      1, 
      0
      ),
  ) %>%
  arrange(RAC1P) %>%
  group_by(SERIALNO) %>%
  summarize_all(first)



# B08105A (means of transport)
# B17020A Poverty 
# B19001A HINCP 
# C27001A Health 
```


LOGIT MODEL

```{r}
sff_pums_clean <- sff_pums_clean %>% 
  mutate(
    race = RAC1P %>% 
      factor(
        levels = sff_pums_clean$RAC1P %>% 
          unique() %>%
          as.numeric() %>% 
          sort()
        ),
    health_coverage = HICOV %>% 
      factor(
        levels = sff_pums_clean$HICOV %>% 
          unique() %>%
          as.numeric() %>% 
          sort()
        ),
    income = HINCP %>% 
      factor(
        levels = sff_pums_clean$HINCP %>% 
          unique() %>%
          as.numeric() %>% 
          sort()
        ),
     PUMA = as.factor(PUMA)
  )

```

```{r}
logit_model <- glm(
  firerisk ~ race + health_coverage + Veh + PUMA,
  family = quasibinomial(),
  data = sff_pums_clean
)

summary(logit_model)

```

